---
title: "ch2"
author: "Francesco Zucca"
format: html
editor: visual
---

# 1 - DATA VISUALIZATION

## 1.1 - Introduction

This chapter covers the visualization of data using the library `ggplot2`\`, which is included in the `tidyverse`\` package. Hence, the first thing to do is to load the `tidyverse`\` package.

```{r}
# loading tidyverse
library("tidyverse")
```

There are some conflicts but we can ignore them safely.

## 1.1.1 - Loading a data frame

To be able to generate any plot, we first need to load a dataset. The book uses the *penguins* dataset, but we will use the *iris* dataset.

```{r}
# loading a dataframe
iris
```

#### The `?`\` operator

The `?`\` operator or the `help()`\` function can be used to access documentation on anything loaded into the R session.

```{r}
?iris
```

We will focus on three main variables: *Sepal.Length*, *Petal.Length*, and *Species*.

## 1.2 - Plotting with `ggplot2`\`

With `ggplot2`\`, you begin a plot with the function `ggplot()`\`, which defines the plot object. Then, you apply layers to it. Hence:

1.  `ggplot()`\`: plot function, with the following main agruments: - `data =`\`: argument that defines the dataset to use in the graph.

    -   `mapping = aes()`\`: argument that defines how variables in your dataset are mapped to visual properties of your plot. The `x`\` and `y`\` arguments of the `aes`\` function specify which variables to map to the x and y axes.

2.  `geom()`\`: function which defines the geometrical object that the plot uses to represent data (note that this is added with a `+`\` after the `ggplot()`\` function.

```{r}
# plotting mpg and hp
ggplot(
  data = iris,
  mapping = aes(x = Sepal.Length, y = Petal.Length) 
) +
  geom_point()
```

### 1.2.1 - Adding aesthetics and layers

#### Adding aesthetics

Different aesthetics can be added as an argument to the `aes`\` function. For example, you can add arguments to give different colors and shapes to different species.

```{r}
# adding color based on species
ggplot(
  data = iris,
  mapping = aes(x = Sepal.Length, y = Petal.Length, 
                color = Species, shape = Species) 
) +
  geom_point()
```

#### Adding layers

More layers can also be added to the plot. For example, you can plot a smooth curve displaying the relationship between the variable on the y axis and the one on the x axis.

```{r}
# adding a smooth curve
ggplot(
  data = iris,
  mapping = aes(x = Sepal.Length, y = Petal.Length, 
                color = Species, shape = Species) 
) +
  geom_point() +
  geom_smooth(method = "lm") # the method specifies the model
```

Notice that if you define aesthetic mappings directly into the `ggplot()`\` function, they get defined as the global level, hence why the `geom_smooth`\` layer is fitting a line for each one of the groups. This can be fixed by not defining the color entry at the global level, but rather in the `geom_point`\` layer (hence at the local level).

```{r}
# fitting only one line
ggplot(
  data = iris,
  mapping = aes(x = Sepal.Length, y = Petal.Length) 
) +
  geom_point(mapping = aes(color = Species, shape = Species)) +
  geom_smooth(method = "lm")
```

### 1.2.2 - ggplot2 calls

When calling a function, argument names can be omitted, and R will automatically assign the values based on their order. For example, the first argument in the `ggplot()`\` function will always be the `data`\` argument.

```{r}
# rewriting the same function but omitting arguments
ggplot(
  iris, aes(Sepal.Length, Petal.Length)
) +
  geom_point(aes(color = Species, shape = Species)) +
  geom_smooth(method = "lm")
```
