---
title: "ch2"
author: "Francesco Zucca"
format: html
editor: visual
---

# 1 - DATA VISUALIZATION

## 1.1 - Introduction

This chapter covers the visualization of data using the library `ggplot2`\`, which is included in the `tidyverse`\` package. Hence, the first thing to do is to load the `tidyverse`\` package.

```{r}
# loading tidyverse
library("tidyverse")
```

There are some conflicts but we can ignore them safely.

## 1.1.1 - Loading a data frame

To be able to generate any plot, we first need to load a dataset. The book uses the *penguins* dataset, but we will use the *iris* dataset.

```{r}
# loading a dataframe
iris
```

#### The `?`\` operator

The `?`\` operator or the `help()`\` function can be used to access documentation on anything loaded into the R session.

```{r}
?iris
```

We will focus on three main variables: *Sepal.Length*, *Petal.Length*, and *Species*.

## 1.2 - Creating a plot with `ggplot2`\`

With `ggplot2`\`, you begin a plot with the function `ggplot()`\`, which defines the plot object. Then, you apply layers to it. Hence:

1.  `ggplot()`\`: plot function, with the following main agruments:
    -   `data =`\`: argument that defines the dataset to use in the graph.

    -   `mapping = aes()`\`: argument that defines how variables in your dataset are mapped to visual properties of your plot. The `x`\` and `y`\` arguments of the `aes`\` function specify which variables to map to the x and y axes.
2.  `geom()`\`: function which defines the geometrical object that the plot uses to represent data (note that this is added with a `+`\` after the `ggplot()`\` function.

```{r}
# plotting mpg and hp
ggplot(
  data = iris,
  mapping = aes(x = Sepal.Length, y = Petal.Length) 
) +
  geom_point()
```
